---
date: "2023-07-05"
title: How Functional is your R code? Introduction to writing your own functions in R
author: "Gary Truong"
---

# Introduction


## R Data Types

Data structures <- Data Types <- Elements (smallest unit possible)


Some common R data types

- Character
- Numeric
- Integer
- Complex Numbers
- Logical
- Factor


Characters (can be letters, numbers, symbols)
```{r}
# using " " will generate character data types
a <- c("a","b","c","d") # letters as characters
class(a)
str(a)

a1 <- c("1","2","3","4") # numbers as characters
str(a1)

a2 <- c("@","#","$","%") # symbols as characters
str(a2)
```

Numeric (decimal numbers)
```{r}
b <- c(1,2,3,4) # numeric class even though they are integers
class(b)
str(b)

b1 <- c(1.1, 2.1, 3.1, 4.1) # numeric decimals
str(b1)
```

Integer (whole numbers)
```{r}
c <- c(1L, 2L, 3L, 4L) # adding the L after the number will make it into integer class
class(c)
str(c)

c1 <- c(1:4) # the colon between the two numbers coerces into integer
str(c1)

# What do we get When we compare the numeric vector with the integer vector
identical(b, c)

identical(c, c1)
```

Complex Numbers
```{r}
d <- c(1i, 2i, 3i, 4i) # complex numbers
class(d)
str(d)
```

Logical
```{r}
e <- c(TRUE, TRUE, FALSE, FALSE) # logical class,
class(e)
str(e)

c(T, T, F, F) # T and F are equivalent to TRUE and FALSE
```

Factor
```{r}
f <- factor(c("Jan", "Feb", "Mar", "Apr")) # creating a factor vector of months with 4 levels
class(f)
str(f)
```

Are there any questions so far? 


## R data structures

- Vectors
- Matrix
- Data Frames
- Lists

Data structures <- Data Types <- Elements


Vectors

Vectors are the simplest data structures in R and probably the most common 

Vectors can only contain elements of the same data type. (i.e. character(a,b,c,d), 
numeric(1,2,3,4), etc)
 
To create a vector in R, use the concatentate (combine) function c()

All of the objects created above are vectors



Question: What happens if we try to combine different data types? 
```{r}
g <- c(1, 2L, "c", TRUE) # what class would this be?
class(g)
str(g)

```


Indexing is used to access individual elements in a data structure

Vector Indexing
```{r}
# Positional Indexing
g <- c(1, 2L, "c", TRUE)

g[1]
g[3:4] 

# logical indexing
g[g == 3] # instead of specifying a position, we use == to match with a value
g[g > "1"] 
```

Matrices and Arrays


```{r}
# 1 dimensional matrices

x <- matrix(1:24, 4, 6)
x[,1]

x1 <- matrix(letters[1:24], 4, 6) 
x1[,1]

# Multidimensional arrays

y <- array(c(x,x1), dim = c(4,6,2)) 
y[,,1]
y[,1,1]
y[,,2]
y[,1,2]
y[,1,]
```


Data Frames and tibbles

A tibble is a data frame but has been altered to work better with the tidyverse packages
```{r}
h <- data.frame(a,b,c,d,e,f,g) # read.csv() to read in data as a data frame
class(h)
str(h)

h1 <- tidyr::tibble(a,b,c,d,e,f,g) # read_csv() to read in data as a tibble
h1
class(h1)
str(h1)

# Indexing a data frame
h
h[1:3,1]
h$a
h$a[1]
h$a[3:4]
h$a[a == "d"]
h[,1]
# h[,,2] = h1
```


List 

Data structure <- Data Types <- Elements

```{r}
i <- list(a,b,c,d,e,f,g,h)
class(i)
str(i)

#indexing a vector within a list
i
i[1]
i[[1]][1]
i[[1]][3:4]

# indexing a data frame within a list
i[[8]]
i[[8]][1]
i[[8]]$a[1]
i[[8]]$a[c(1,3:4)]

#You can also have a list within a list 
```


